### 2024-10-26 report/sections/introduction.Rmd
After review update

First, writers and speakers can in principle plan utterances in full for one or multiple sentences before beginning to output words. However this is not what people typically do: utterances are not fully formed at production onset. Instead, syntax and lexical content and even details of the message itself are planned as the emergent result of a real-time process. For speech it is well known that the pre-planning duration depends -- among other factors -- on the need to avoid intra-sentential hesitations; speakers have a preference towards minimising language pre-planning depending on the extent to which upcoming words can be planned in parallel to production [e.g. @levelt2000word; @griffin2003reversed; @wheeldon_konopka_2023]. For writing, there is evidence that the time to keystroke / pen onset increases when a sentence starts with a more complex sentence-initial phrases similar as in speech [@not10; dam09; @roeser2018advance]. For example, in @roeser2018advance participants described moving arrays of images in simple utterances such as *A and the B moved above the C* that either started with a conjoined noun phrase (i.e. *A and the B moved*) or a simple noun (i.e., *A moved above the B and the C*). Importantly while the overall complexity of the utterance (in terms of length, number of phrases and noun) was held constant, sentences that started with a conjoined noun phrase increased the time to typing onset (and speech onset). This is not what one would expect if writers plan sentences in full before production onset. In fact, eye-movement data reported in @torrance2012written and @roeser2018advance demonstrate that writers do not plan the lexical form of the utterance beyond the first noun before production onset [similar to speech @gri01; @griffin2003reversed], although there was evidence that lexical pre-planning of the second noun (i.e. *the B*) was more likely for conjoined noun phrases, i.e. when the second noun was part of the same phrase as the first noun.

>>>>

First, writers and speakers can in principle plan utterances in full for one or multiple sentences before beginning to output words. However this is not what people typically do: utterances are not fully formed at production onset. Instead, syntax and lexical content and even details of the message itself are planned as the emergent result of a real-time process. For speech it is well known that the pre-planning duration depends -- among other factors -- on the need to avoid intra-sentential hesitations; speakers have a preference towards minimising language pre-planning depending on the extent to which upcoming words can be planned in parallel to production [e.g. @levelt2000word; @griffin2003reversed; @wheeldon_konopka_2023]. For writing, there is evidence that the time to keystroke / pen onset increases when a sentence starts with a more complex sentence-initial phrases similar as in speech [@not10; @dam09; @roeser2018advance]. For example, in @roeser2018advance participants described moving arrays of images in simple utterances such as *A and the B moved above the C* that either started with a conjoined noun phrase (i.e. *A and the B moved*) or a simple noun (i.e., *A moved above the B and the C*). Importantly while the overall complexity of the utterance (in terms of length, number of phrases and noun) was held constant, sentences that started with a conjoined noun phrase increased the time to typing onset (and speech onset). This is not what one would expect if writers plan sentences in full before production onset. In fact, eye-movement data reported in @torrance2012written and @roeser2018advance demonstrate that writers do not plan the lexical form of the utterance beyond the first noun before production onset [similar to speech @gri01; @griffin2003reversed], although there was evidence that lexical pre-planning of the second noun (i.e. *the B*) was more likely for conjoined noun phrases, i.e. when the second noun was part of the same phrase as the first noun.

### 2024-10-26 report/sections/methods.Rmd
After review update

```{r datasets, results='asis'}
tibble(Dataset = c("C2L1", "CATO", "SPL2", "PLanTra", "LIFT", "GE2"),
                Source = c("Rønneberg et al. (2022)",
                           "Torrance et al. (2016)",
                           "Torrance et al. (n.d.)",
                           "Rossetti and Van Waes (2022b)",
                           "Vandermeulen, Steendam, et al. (2020)", 
                           "Torrance and Ofstad (n.d)"),
                Keylogger = c("EyeWrite", 
                           "EyeWrite", 
                           "CyWrite", 
                           "InputLog", 
                           "InputLog",
                           "EyeWrite"),
                `Writing task` = c("Argumentative", 
                                 "Expository",
                                 "Argumentative",
                                 "Text simplification",
                                 "Synthesis", 
                                 "Argumentative"),
                 N = as.character(c(126, 26*2, 39, 47, 658, 45)),
#                n_texts = c(1, 2, 2, 2, NA),
#                n_sentences = c(),
#                n_words = c(),
                `Mean age` = c(11.8, 16.9, 20.6, 23, 17.0, 19),
                 Sample = c("6th graders", "Secondary school students (dyslexic, non dyslexic)", "Undergraduate students", "Master students", "Secondary school students", "Undergraduate students"),
                 Country = c("Norway", "Norway", "USA", "Belgium", "The Netherlands", "UK"),
                 Language = c( "Norwegian", 
                               "Norwegian", 
                               "English", 
                               "English (L2)", 
                               "Dutch",
                               "English")) %>% 
  arrange(Dataset) %>% 
  apa_table(align = c("l", "p{3cm}", "p{2cm}", "p{2.5cm}", "r", "r", "p{2.5cm}", "p{2.25cm}", "p{2.5cm}"), 
            escape = FALSE, 
            font_size = "footnotesize",
            caption = "Datasets in brief.")

#                Manipulation = c("--", 
#                               "Weak decoders / control; texted masked / unmasked",
#                               "Text produced in L1 / L2",
#                               "Pre / post test with training on plain language principles", 
#                              "Various topics and genres",
#                               "masked / unmasked")
```

@@

From the keystroke data we extracted interkey intervals between adjacent keys at locations that have been identified repeatedly in previous research as locations where observed mean interkey intervals (or counts of interkey intervals over a predetermined threshold) have tended to vary substantially [e.g. @chukharev2019combined; @torrance2016adolescent; @de2018exploring] and are detailed in Table \ref{tab:keyloc}. In particular we analysed the interkey intervals that resulted in the insertion of a character that started a new sentence (before-sentence transition); interkey intervals that started a new word (other than those at the beginning of a sentence;  before-word transitions); and transitions between characters within words all transitions. At before-sentence locations, interkey intervals were timed to the shift keypress that resulted in the capitalization of the first key for most data sets (CATO, C2L1, SPL2, GE2) but were timed to the character key press in the  PLanTra and LIFT datasets (i.e. the duration included the time to perform the capitalisting shift keypress); we return to this difference in the Results section. Transitions that occurred at the very start of the text or at the beginning of a paragraph were not treated as before-sentence transitions and were removed from the analysis. Importantly we also removed transitions that were followed by an editing operation. We therefore just modelled the times between keypresses in ongoing production.

@@

table <- tibble(`Transition type` = c("Within word", "Below word", "Before sentence"),
       Description = c("Transitions between any letter",
                       "Keypress after space followed by any letter",
                       "Keypress following a space preceding any letter"),
       Example = c("T$^{\\wedge}$h$^{\\wedge}$e c$^{\\wedge}$a$^{\\wedge}$t m$^{\\wedge}$e$^{\\wedge}$o$^{\\wedge}$w$^{\\wedge}$e$^{\\wedge}$d. T$^{\\wedge}$h$^{\\wedge}$a$^{\\wedge}$t[bsp][bsp]e$^{\\wedge}$n i$^{\\wedge}$t s$^{\\wedge}$l$^{\\wedge}$e$^{\\wedge}$p$^{\\wedge}$t.", 
                   "The $^{\\wedge}$cat $^{\\wedge}$meowed. That[bsp][bsp]en $^{\\wedge}$it $^{\\wedge}$slept.", 
                   "The cat meowed. $^{\\wedge}$That[bsp][bsp]en it slept."))

apa_table(table, caption = "Transition location classification.",
            align = c("p{3cm}", "p{5cm}", "p{6cm}"), 
            escape = FALSE, 
            font_size = "footnotesize",
            note = "$'^{\\wedge}$' marks transition location; [bsp] represents backspace.")

@@

apa_table(datareduction, 
          escape = FALSE,  
          placement = "bp!",
          caption = "Data reduction. Mean percentage of extreme data removed and the mean percentage of randomly sampled data analysed by transition location (`--` if all data were used). Standard error in parentheses.",
          align =c("l", rep("r", 5)),
          col_spanners = list(" " = 1,
                              "Extreme values in \\%" = c(2, 3),
                              "Data analysed in \\%" = c(4, 6)))

@@

```{r eval=F}
models <- tibble(Models = str_c("M",1:5),
       Type = c("LMM", "LMM", "LMM", "MoG", "MoG"),
       `Eq.` = str_c("\ \\ref{eq:", c("unimodgaus", "unimodloggaus", "unimoduv", "bimodcon", "bimoduncon"), "} "),
#       group = c(rep("Serial", 3), rep("Cascading", 2)),
       Description = c("Single Gaussian model with effects of transition location.",
                       "Single distribution log-Gaussian model equivialent to M1.",
                       "Equivalent to M2 but with different variance components for each transition location (unequal variance).",
                       "Two-distributions mixture of a log-Gaussian for fluent interkey intervals and another wider log-Gaussian for hesitant / disfluent interkey intervals; the distribution of hesitant interkey intervals assumes -- similar to M3 -- different values for transition locations. The mixing weight of these distributions capture the relative number of disfluent transitions for each transition location.",
                       "Equivialent to M4 but distribution of fluent transitions is allowed to vary by transition location.")) %>% 
  select(Models, Description)
```

>>>>

```{r datasets, results='asis'}
tibble(Dataset = c("C2L1", "CATO", "SPL2", "PLanTra", "LIFT", "GE2"),
                Source = c("Rønneberg et al. (2022)",
                           "Torrance et al. (2016)",
                           "Torrance et al. (n.d.)",
                           "Rossetti and Van Waes (2022b)",
                           "Vandermeulen, Steendam, et al. (2020)", 
                           "Torrance and Ofstad (n.d)"),
                Keylogger = c("EyeWrite", 
                           "EyeWrite", 
                           "CyWrite", 
                           "InputLog", 
                           "InputLog",
                           "EyeWrite"),
                `Writing task` = c("Argumentative essay", 
                                 "Expository essay",
                                 "Argumentative essay",
                                 "Paraphrase technical text$^b$",
                                 "Synthesis of several printed sources$^b$", 
                                 "Argumentative essay"),
                 `Words in text$^a$` = c("128.5 (52.9)", "288 (119)", "668 (301)", "295 (79)", "374 (56)$^d$", "540 (141)"),
#                n_texts = c(1, 2, 2, 2, NA),
#                n_sentences = c(),
#                n_words = c(),
                `Age` = c("11.8 (0.47)", "16.9 (0.88)", "20.6 (1.62)", "23 (2.7)", "16.7 (1.5)", "19.1 (1.4)$^c$"),
                 N = as.character(c(126, 26*2, 39, 47, 658, 45)),
                 Sample = c("6th graders", "Secondary school students (dyslexic, non dyslexic)", "Undergraduate students", "Master students", "Secondary school students", "Undergraduate students"),
                 Country = c("Norway", "Norway", "USA", "Belgium", "The Netherlands", "UK"),
                 Language = c( "Norwegian", 
                               "Norwegian", 
                               "English", 
                               "English (L2)", 
                               "Dutch",
                               "English")) %>% 
  arrange(Dataset) %>% 
  apa_table(align = c("p{1.5cm}", "p{2.75cm}", "p{1.75cm}", "p{2.25cm}", "p{1.75cm}", "p{1.75cm}", "p{.5cm}", "p{2.5cm}", "p{1.75cm}", "p{2cm}"), 
            escape = FALSE, 
            longtable = FALSE,
            row.names = T,
            font_size = "footnotesize",
            caption = 'Datasets in brief.', 
            note = "$^a$Mean number of words in the final text; standard deviations in parentheses. $^b$Reading activity was marked in the dataset and removed before analysis. $^c$Estimated from a similar sample from the same population. $^d$Estimated based on the number of characters in the final product following Brysbaert et al. (2021).")

@@

From the keystroke data we extracted interkey intervals between adjacent keys at locations that have been identified repeatedly in previous research as locations where observed mean interkey intervals (or counts of interkey intervals over a predetermined threshold) have tended to vary substantially [e.g. @chukharev2019combined; @torrance2016adolescent; @de2018exploring] and are detailed in Table \ref{tab:keyloc}. In particular we analysed the interkey intervals that resulted in the insertion of a character that started a new sentence (before-sentence transition); interkey intervals that started a new word (other than those at the beginning of a sentence;  before-word transitions); and transitions between characters within words all transitions. At before-sentence locations, interkey intervals were timed to the shift keypress that resulted in the capitalization of the first key for most data sets (CATO, C2L1, SPL2, GE2) but were timed to the character key press in the  PLanTra and LIFT datasets (i.e. the duration included the time to perform the capitalising shift keypress); we return to this difference in the Results section. Transitions that occurred at the very start of the text or at the beginning of a paragraph were not treated as before-sentence transitions and were removed from the analysis. Importantly we also removed transitions that were followed by an editing operation. We therefore just modelled the times between keypresses in ongoing production.

@@

table <- tibble(`Transition type` = c("Within word", "Before word", "Before sentence"),
       Description = c("Transitions between any letter",
                       "Keypress after space followed by any letter",
                       "Keypress following a space preceding any letter"),
       Example = c("T$^{\\wedge}$h$^{\\wedge}$e c$^{\\wedge}$a$^{\\wedge}$t m$^{\\wedge}$e$^{\\wedge}$o$^{\\wedge}$w$^{\\wedge}$e$^{\\wedge}$d. T$^{\\wedge}$h$^{\\wedge}$a$^{\\wedge}$t[bsp][bsp]e$^{\\wedge}$n i$^{\\wedge}$t s$^{\\wedge}$l$^{\\wedge}$e$^{\\wedge}$p$^{\\wedge}$t.", 
                   "The $^{\\wedge}$cat $^{\\wedge}$meowed. That[bsp][bsp]en $^{\\wedge}$it $^{\\wedge}$slept.", 
                   "The cat meowed. $^{\\wedge}$That[bsp][bsp]en it slept."))

apa_table(table, caption = 'Transition location classification. The final text in all cases was "The cat meowed. Then it slept."',
            align = c("p{3cm}", "p{5cm}", "p{6cm}"), 
            escape = FALSE, 
            font_size = "footnotesize",
            note = "$'^{\\wedge}$' marks transition location; [bsp] represents backspace.")

@@

apa_table(datareduction, 
          escape = FALSE,  
          placement = "bp!",
          caption = "Data reduction. Shown are the percentage of extreme data that were removed and the percentage of randomly sampled data that were analysed by transition location (`--` if all data were included). Standard error in parentheses.",
          align =c("l", rep("r", 5)),
          col_spanners = list(" " = 1,
                              "Extreme values in \\%" = c(2, 3),
                              "Data analysed in \\%" = c(4, 6)))

@@

```{r eval=F}
models <- tibble(Models = str_c("M",1:5),
       Type = c("LMM", "LMM", "LMM", "MoG", "MoG"),
       `Eq.` = str_c("\ \\ref{eq:", c("unimodgaus", "unimodloggaus", "unimoduv", "bimodcon", "bimoduncon"), "} "),
#       group = c(rep("Serial", 3), rep("Cascading", 2)),
       Description = c("Single Gaussian model with effects of transition location.",
                       "Single distribution log-Gaussian model equivialent to M1.",
                       "Equivalent to M2 but with different variance components for each transition location (unequal variance).",
                       "Two-distributions mixture of a log-Gaussian for fluent interkey intervals and another wider log-Gaussian for hesitant interkey intervals; the distribution of hesitant interkey intervals assumes -- similar to M3 -- different values for transition locations. The mixing weight (i.e. proportion of observations estimated to be associated with each distribution) captures the relative number of hesitant transitions for each transition location.",
                       "Equivialent to M4 but distribution of fluent transitions is allowed to vary by transition location.")) %>% 
  select(Models, Description)
```

### 2024-10-26 report/sections/results.Rmd
After review update

```{r loos, results='asis'}
table %>% 
  arrange(desc(Comparison)) %>% 
  pivot_longer(-Comparison, names_to = "Dataset") %>% 
  pivot_wider(names_from = Comparison, values_from = value) %>% 
  relocate(Dataset, ends_with("M4)")) %>% 
  arrange(Dataset) %>% 
  apa_table(caption = "Model comparisons. Models were compared incrementally from the simplest to the most complex model. Comparison are shown with the distribution type of the model with the higher predictive performance. A negative difference in $\\Delta\\widehat{elpd}$ indicates higher predictive performance for the more complex model; standard error shown in parentheses.",
            align =c("p{1.75cm}", "p{4cm}", "p{4.25cm}", "p{3.5cm}"), 
            escape = FALSE,
            row.names = T,
            col.names = c("Dataset", "Mixture of two log-Gaussians (M3 -- M4)", "Single log-Gaussian (unequal variance; M2 -- M3)", "Single log-Gaussian (M1 -- M2)"),
            longtable = FALSE,
            font_size = "small",
            note = "$\\Delta\\widehat{elpd}$ = difference in predictive performance -- estimated as expected log pointwise predictive density")

@@

# add axes titles
grid.arrange(patchworkGrob(plots_fin), 
                          left = "Density",
                          bottom = "Interkey interval in secs")
#ggsave("mixdists.png", plots_fin, width = 12, height = 6)

@@

ps <- purrr::map_dfr(files, ~read_csv(.x) %>% 
  mutate(data = .x)) %>% 
  mutate(across(data, ~sub(".*/([^/]+)/[^/]+\\..*", "\\1", .)),
         across(data, ~recode_factor(., 
                "c2l1" = "C2L1",
                "cato" = "CATO",# (non-dyslexic, unmasked)",
                "gunnexp2" = "GE2",# (unmasked)",
                "lift" = "LIFT",
                "plantra" = "PLanTra",
                "spl2" = "SPL2", # (L1)", 
#                "spl2 (shift + C)" = "SPL2 (L1; shift + C)", 
                .ordered = TRUE))) %>% 
  filter(!(str_detect(data, "SPL2") & lang == "ES" & !is.na(lang)),
         !(str_detect(data, "CATO") & group == "dyslexic" & !is.na(group)),
         !(str_detect(data, "CATO") & task == "masked" & !is.na(task)),
         !(str_detect(data, "GE2") & xn == "masked" & !is.na(xn)))

@@

```{r crossstudypost2, fig.cap="Mixture model parameter estimates across studies. Distributions of parameter estimates are represented as posterior mean and 95\\% probability interval (PI). Estimates for the CATO dataset were calculated for the non-dyslexic group, unmasked condition; also the GE2 estimtes represent the unmasked condition; SPL2 estimates are for the L1 group."}

@@

grid.arrange(patchworkGrob(plot), 
             left = "Posterior estimate with 95% PIs",
             bottom = "Transition location")

@@

@@

apa_table(tmp,
          align = c("l", rep("p{3cm}", 4)), 
#          font_size = "small",
          longtable = FALSE,
          escape = FALSE,
          row.names = T,
          col.names = c("Dataset", rep(c("before sentence vs word", "before vs within word"), 2)),
          col_spanners = list(" " = c(1),
                              "Hesitation slowdown" = c(2, 3), 
                              "Hesitation probability" = c(4, 5)),
          note = "BF = evidence in favour of the alternative hypothesis over the null hypothesis.",
          caption = "BFs for the effect of transition location on interkey intervals shown for the hesitation slowdown and hesitation probability. Comparisons for the CATO dataset were calculated for the non-dyslexic group, unmasked condition; also the GE2 estimtes represent the unmasked condition; SPL2 estimates are for the L1 group.")

>>>>

```{r loos, results='asis'}
table %>% 
  arrange(desc(Comparison)) %>% 
  pivot_longer(-Comparison, names_to = "Dataset") %>% 
  pivot_wider(names_from = Comparison, values_from = value) %>% 
  relocate(Dataset, ends_with("M4)")) %>% 
  arrange(Dataset) %>% 
  apa_table(caption = "Model comparisons. Models were compared incrementally from the simplest to the most complex model. Comparison are shown with the distribution type of the model with the higher predictive performance. A negative difference in $\\Delta\\widehat{elpd}$ indicates higher predictive performance for the more complex model; standard error shown in parentheses.",
            align =c("p{2cm}", "p{4cm}", "p{4.25cm}", "p{3.5cm}"), 
            escape = FALSE,
            row.names = T,
            col.names = c("Dataset", "Mixture of two log-Gaussians (M3 -- M4)", "Single log-Gaussian (unequal variance; M2 -- M3)", "Single log-Gaussian (M1 -- M2)"),
            longtable = FALSE,
            font_size = "small",
            note = "$\\Delta\\widehat{elpd}$ = difference in predictive performance -- estimated as expected log pointwise predictive density")

@@

# add axes titles
grid.arrange(patchworkGrob(plots_fin), 
                          left = "Density",
                          bottom = "Interkey intervals in secs")
#ggsave("mixdists.png", plots_fin, width = 12, height = 6)

@@

ps <- purrr::map_dfr(files, ~read_csv(.x) %>% 
  mutate(data = .x)) %>% 
  mutate(across(data, ~sub(".*/([^/]+)/[^/]+\\..*", "\\1", .)),
         across(data, ~recode_factor(., 
                "c2l1" = "C2L1",
                "cato" = "CATO",# (non-dyslexic, unmasked)",
                "gunnexp2" = "GE2",# (unmasked)",
                "lift" = "LIFT",
                "plantra" = "PLanTra",
                "spl2" = "SPL2", # (L1)", 
#                "spl2 (shift + C)" = "SPL2 (L1; shift + C)", 
                .ordered = TRUE))) %>% 
  filter(!(str_detect(data, "SPL2") & lang == "ES" & !is.na(lang)),
         !(str_detect(data, "CATO") & group == "dyslexic" & !is.na(group)),
         !(str_detect(data, "CATO") & task == "masked" & !is.na(task)),
         !(str_detect(data, "GE2") & xn == "masked" & !is.na(xn))) %>% 
  mutate(across(data, ~factor(., levels = unique(data)[6:1], ordered = T)),
         across(location, factor))

@@

```{r crossstudypost2old, fig.cap="Mixture model parameter estimates across studies. Distributions of parameter estimates are represented as posterior mean and 95\\% probability interval (PI). Estimates for the CATO dataset were calculated for the non-dyslexic group, unmasked condition; also the GE2 estimtes represent the unmasked condition; SPL2 estimates are for the L1 group.", eval = F}

@@

grid.arrange(patchworkGrob(plot), 
             left = "Posterior estimates with 95% PIs",
             bottom = "Transition location")

@@

```{r crossstudypost2, fig.cap="Mixture model parameter estimates across studies. Distributions of parameter estimates are represented as posterior mean and 95\\% probability interval (PI). Interkey-interval duration measures are shown in msecs and hesitation probability is shown in proportions. Estimates for the CATO dataset were calculated for the non-dyslexic group, unmasked condition; also the GE2 estimtes represent the unmasked condition; SPL2 estimates are for the L1 group.", fig.height=5.5}

posd <- position_dodge(-.5)
dotsize <- 2.5
grouplabel <- "Transition\nlocations:"
shapes <- c(8, 9, 1)
width <- 0
linewidth <- .35
nrow <- 1
beta <- filter(ps, name == "beta", 
               location == "before\nword") %>%
  mutate(name = beta_label,
         location = "overall") %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = data)) +
  geom_errorbar(width = width, alpha = .75) +
  geom_point(size = dotsize) +
  facet_grid(~name) +
  scale_shape_manual(values = shapes) +
  scale_colour_colorblind() +
  scale_y_continuous(labels = scales::comma) +
  theme(axis.title = element_blank(),
        axis.ticks.y = element_blank()) +
  coord_flip() +
  labs(colour = grouplabel,
       shape = grouplabel) +
  guides(colour=guide_legend(nrow=nrow,byrow=TRUE),
         shape = guide_legend(nrow=nrow,byrow=TRUE))

delta <- filter(ps, name == "beta2") %>% 
  mutate(name = delta_label) %>% 
  #mutate(across(c(mean, lower, upper), ~./1000)) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = data,
             shape = location)) +
  geom_errorbar(width = width, alpha = .75, position = posd) +
  geom_point(aes(colour = location), 
            size = dotsize, 
            position = posd) +
  facet_grid(~name) +
  scale_shape_manual(values = shapes) +
  scale_colour_colorblind() +
  scale_y_continuous(labels = scales::comma) +
  labs(colour = grouplabel,
       shape = grouplabel) +
  theme(axis.title = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  coord_flip() +
  guides(colour=guide_legend(nrow=nrow,byrow=TRUE),
         shape = guide_legend(nrow=nrow,byrow=TRUE))

theta <- filter(ps, name == "prob") %>% 
  mutate(name = theta_label) %>% 
  ggplot(aes(y = mean, 
             ymin = lower, 
             ymax = upper,
             x = data,
             shape = location)) +
  geom_errorbar(width = width, alpha = .75, position = posd) +
  geom_point(aes(colour = location), 
            size = dotsize, 
            position = posd) +
  facet_grid(~name) +
  scale_shape_manual(values = shapes) +
  scale_colour_colorblind() +
  scale_y_continuous(labels = dezero_plot, limits = c(0, 1)) +
  labs(colour = grouplabel,
       shape = grouplabel) +
  theme(axis.title = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) + 
  coord_flip() +
  guides(colour=guide_legend(nrow=nrow,byrow=TRUE),
         shape = guide_legend(nrow=nrow,byrow=TRUE))

plot <- beta + delta + theta + plot_layout(guides = "collect",
                                           widths = c(.4, 1, 1))

plot <- grid.arrange(patchworkGrob(plot), 
             bottom = "Posterior estimates with 95% PIs")

ggsave("../posteriorplot.pdf", plot = plot, width = 8, height = 6)
```

@@

apa_table(tmp,
          align = c("l", rep("p{3cm}", 4)), 
#          font_size = "small",
          longtable = FALSE,
          escape = FALSE,
          row.names = T,
          col.names = c("Dataset", rep(c("before sentence vs word", "before vs within word"), 2)),
          col_spanners = list("Hesitation duration" = 2:3, 
                              "Hesitation probability" = 4:5),
          note = "BF = evidence in favour of the alternative hypothesis over the null hypothesis calculated using the Savage-Dickey method (Dickey et al., 1970; Wagenmakers et al., 2010). BFs greater than 3 indicate moderate evidence, greater than 10 strong evidence, and greater than 30 very strong evidence for a statistically meaningful effect (Jeffreys, 1961; Lee \\& Wagenmakers, 2014). A BF smaller than 0.33 suggests evidence in favour of a null effect (Dienes, 2014) and a BF around 1 is inconclusive evidence.",
          caption = "BFs for the effect of transition location on interkey intervals shown for the hesitation duration and hesitation probability. Comparisons for the CATO dataset were calculated for the non-dyslexic group, unmasked condition; also the GE2 estimtes represent the unmasked condition; SPL2 estimates are for the L1 group.")

### 2024-10-26 slides/sections/discussion.Rmd
After review update

- Writers do not necessarily pause but plan utterances in parallel to writing.
- Evidenced by 
  - stronger predictive performance of mixture models [@roeser2021modelling].
  - pauses are generally equally likely before before sentences compare to before words (but often longer).
- Largely consistent for populations with different levels of writing experience, languages (e.g. young / L2 writers, students), composition tasks. 
- First robust evidence that written composition is largely a parallel process.

>>>>

- First robust evidence that written composition is largely a parallel process.
- Writers do not necessarily pause at larger linguistic locations but plan utterances in parallel to writing.
- Evidenced by 
  - stronger predictive performance of mixture models [@roeser2021modelling].
  - pauses are not generally more likely before sentences compare to words (but often longer).
- Pauses are consistently more likely at word-initial location compared to word-medial but (1) not always longer and (2) for sentence-initial interkey intervals pausing behaviour varies as a function of writing experience, languages (e.g. young / L2 writers, students), composition tasks.

### 2024-10-26 slides/sections/evidence.Rmd
After review update

- Interkey intervals are longer at larger linguistic edges following relatively fluent production bursts [@chenoweth2001fluency; @hayes2012evidence; @kaufer1986composing].
- Mean pause duration before a writer starts a sentence is typically longer than before a mid-sentence word and these are, in turn, longer than between mid-word key presses [e.g. @conijn2019understanding]. 
- Planning message and syntax for the next sentence adds to the time required to prepare the upcoming word, and the motor planning required to produce the first keystroke [@baaijen2012keystroke; @roeser2018advance].

@@

- Writers often do not hesitate before sentence (or word) onset.
- In some cases sentences are not planned, planning was postponed until after sentence onset, or planning was completed in parallel with previous output [@not10; dam09; @roeser2018advance].
- Sentence initial durations are often shorter than what one would expect if sentence-initial pauses reflect planning.
  - @medimorec2017pauses: 71% of sentence-initial interkey intervals shorter than 1 sec. 
  - @ronneberg2022process: 50% sentence-initial intervals were around 430 msecs but 1.2 secs for the remainder. 
  <!-- - For comparison, these short pauses are less than mean written naming response latency for single objects [@torrance2018timed] and     longer pauses were similar to response latencies for short sentences when describing arrays of images [@roeser2018advance].  -->
- Fluent output is maintained by a cascade of processes that run in parallel [@olive2014toward;@van1991handwriting].

## Two distinct data-generating processes leading to a mixture of two distributions of interkey intervals [@roeser2021modelling].

- Minimum interkey intervals for executing motor movements; 100-150 msecs [@conijn2019understanding;@waes2019].
- If upstream processes provide output more slowly, then interkey intervals are determined by time taken to complete upstream processes and not on finger movement. 
- Neither process (or distribution of keystrokes) can be explained on the basis of the keystroke location in the text.

>>>>

- Pause duration before a writer starts a sentence is typically longer than before a mid-sentence word and these are longer than between mid-word key presses [e.g. @conijn2019understanding]. 
- Long interkey intervals are typically followed by production bursts [@hayes2012evidence; @kaufer1986composing].
- Planning message and syntax for a sentence adds to the time required to prepare the upcoming word, and the motor planning required to produce the first keystroke [@baaijen2012keystroke; @roeser2018advance].

@@

- Variations in interkey intervals are not sufficiently explained by text location.
- Writers often do not hesitate before sentence (or word) onset.
- In some cases sentences are not planned, planning was postponed until after sentence onset, or completed in parallel with previous output [@not10; @roeser2018advance].
- Sentence initial durations are often shorter than what one would expect if sentence-initial pauses reflect planning.
- Fluent output is maintained by a cascade of processes that run in parallel [@olive2014toward; @van1991handwriting].
  <!-- - @medimorec2017pauses: 71% shorter than 1 sec -->
  <!-- - @ronneberg2022process: 50% around 430 msecs and around 1.2 secs  -->
  <!-- - For comparison, these short pauses are less than mean written naming response latency for single objects [@torrance2018timed] and longer pauses were similar to response latencies for short sentences when describing arrays of images [@roeser2018advance].  -->

### 2024-10-26 slides/sections/implementation.Rmd
After review update

## Implentation

- Bayesian inference [e.g. @lambert2018student]
- Random effects for participants [see @roeser2021modelling] with code based on @vasishth2017
- Stan [@carpenter2016stan] using `rstan` [@rstan] to interface with $R$
- Leave-one-out cross-validation estimated as the sum of the expected log predictive density $\widehat{elpd}$ and compared by the difference between models $\Delta\widehat{elpd}$ [@vehtari2015pareto; @vehtari2017practical]
- Difference between models was summarised as $\mid\frac{\Delta\widehat{elpd}}{\text{SE}}\mid$, the *z*-score of the difference between models [@sivula2020uncertainty].

>>>>

## Implementation and model comparison

<!-- - Bayesian inference [e.g. @lambert2018student] models implemented in Stan [@carpenter2016stan] using `rstan` [@rstan] to interface with $R$ -->
- Code based on @vasishth2017 [see @roeser2021modelling]
- Random effects for participants
- Leave-one-out cross-validation: sum of the *expected log predictive density* $\widehat{elpd}$ 
- *Difference between models* $\Delta\widehat{elpd}$ [@vehtari2015pareto; @vehtari2017practical]
- Difference between models was summarised as

$\mid\frac{\Delta\widehat{elpd}}{\text{SE}_\text{diff}}\mid$,

i.e. the *z*-score of the change in predictive performance [@sivula2020uncertainty].
